resistanceModelNotesAndy

### 2016-03-24 paper1_results_figs.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

  #add an extra predictor, the lower starting freq of resistance divided by the larger
  resist_start_hi_div_lo <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
  treeInput <- rbind(treeInput,resist_start_hi_div_lo)

@@

  #create a curtisInputs dataframe for use later
  curtisInputs <- data.frame("exposure"=0.9,
                             "effectiveness_insecticide1"=0.73,
                             "effectiveness_insecticide2"=1,
                             "start_freq_allele1"=0.01,
                             "start_freq_allele2"=0.01,
                             "dominance_allele1"=0.17,
                             "dominance_allele2"=0.0016,
                             "selection_coef_allele1"=0.23,
                             "selection_coef_allele2"=0.43,
                             "maleExposureProp"=1,
                             "correctMixDeployProp"=1,
                             "resist_start_hi_div_lo"=1,
                             "resist_start_1_div_2"=1)

@@

    #modify killer plot to add in male exposure prop on x axis
    library(viridis)
    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+

>>>>

  #add an extra predictor, the lower starting freq of resistance divided by the larger
  resist_start_lo_div_hi <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
  treeInput <- rbind(treeInput,resist_start_lo_div_hi)

@@

  #create a curtisInputs dataframe for use later
  curtisInputs <- data.frame("exposure"=0.9,
                             "effectiveness_insecticide1"=0.73,
                             "effectiveness_insecticide2"=1,
                             "start_freq_allele1"=0.01,
                             "start_freq_allele2"=0.01,
                             "dominance_allele1"=0.17,
                             "dominance_allele2"=0.0016,
                             "selection_coef_allele1"=0.23,
                             "selection_coef_allele2"=0.43,
                             "maleExposureProp"=1,
                             "correctMixDeployProp"=1,
                             "resist_start_lo_div_hi"=1,
                             "resist_start_1_div_2"=1)

@@

    #modify killer plot mix2 vs seq to add in male exposure prop on x axis
    library(viridis)
    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+

### 2016-03-24 paper1_results_figs_slimmed_50.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

Grey triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels). The grey line is a linear model through these same points.

@@

>>>>

Red triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels). The red line is a linear model through these same points.

@@

## Fig. x5.1 modified for mix2 vs seq

fewer runs where sequential gives slower resistance

  ggSubset <- dif_mix2_seq
  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)

  #tricky
  #be a bit careful what to do with zeros where results are same for both strategies
  #I use these better vars to facet the plots
  #may want the zeroes in both facets
  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
  #set these to 1 (they were 0)
  df_same$mix2_better_seq0.5 <- 1
  df_same$seq_better_mix20.5 <- 1
  #rbind them back onto other dataset
  ggSubset <- rbind( ggSubset, df_same)

  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )

    #find unequal runs
    unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
    nrow(ggSubset)
    nrow(unequal)
    #2721 out of the 6189 remaining runs

    #in how many runs was seq better than mix
    sum(unequal$seq_better_mix20.5)
    #311

    library(viridis)

    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+

             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\nresistance slower','1'='sequential\nresistance slower'))) +

             #now add the unequal points
             #showed that they were on both sides
             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=2) +

  l_mod <- lm(y1 ~ x1)
  #l_mod$coefficients
  #(Intercept)          x1 
  #1.019330    0.560508

### 2016-03-24 resistanceModelNotesAndy.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

Can I show this by modifying the fig2 plot ?

@@

1. add option to rescale x axis to years or months in plotcurtis_f2_generic, generation=1month
1. tidy these notes

@@

### current state of play 18/12/15
1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
1. runModel2.r          : runs model for the passed scenarios

>>>>

Can I show this by modifying the fig2 plot ?
Although probably we would need to rerun a sensitivity type anlysis if we wanted to address this properly ...

curtis f2
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )

curtis f2 with i1 set to high starting freq (0.5 breaks plot slightly, but still seems to show that resistance for i2 takes longer in mixture)
runcurtis_f2( P_1 = 0.5 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )

curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
runcurtis_f2( P_1 = 0.45 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )

curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
runcurtis_f2( P_1 = 0.01 , P_2 = 0.45 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )

similar with P_2 0.3
runcurtis_f2( P_1 = 0.01 , P_2 = 0.3 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )

but what about the params for curtis_tab1_ex_iv ?
Beths analysis table uses the same params I think.

As a first step I've added an extra fig x8 onto the end of the figures document attached and in dropbox (I started to address figure tweaks, but have held fire on that while we sort this). Fig x8 addresses whether to keep using an old insecticide when a new one is introduced. It just increases the starting frequency of each resistance in turn to just below the threshold (0.49). I think this accomplishes the same as Beths table 4.

Would I be able to alter the parameters to recreate Beths table3 and Curtis table 1 example vi in a similar way ? I couldn't quite work out which inputs to change.

How would we investigate this in a sensitivity analysis ? Essentially we could repeat what we've done already but with unequal starting frequencies. We had the issue before of not wanting one frequency to be dependent on the other. So we could simply set P_1 to a low range as in current analysis, and then have P_2 at a higher range (e.g. 0.3-0.5). Or perhaps we don't need a range for the 'old' insecticide we could just set at 0.5 ?

I'm slightly nervous about running more analyses but I could investigate this with a couple of days work. Then we would need to decide how to squeeze into the paper.

Happy to chat,

Andy

### previous current state of play 23/3/16
1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
1. runModel2.r          : runs model for the passed scenarios
1. paper1_results_figs_slimmed50.Rmd : final figures for paper

What would I need to do ?

1. sensiAnPaperPart.r : set new param ranges for P_2, perhaps add an argument to enable running them 
1. sensiAnPaper1All.Rmd : rerun calling new version of sensiAnPaperPart, save results to something like : listOutMix_unequal_10000.rda
1. paper1_results_figs_slimmed50.Rmd : save with new name, just rerun using the new results files

perhaps do for 500 runs first ?

reran new unequal scenarios with start_freq set to constant 0.5.
first plots OK, but then get a weird thing where mixa_minus_seq0.5 is always -2. So mix adapative is always taking 2 generations less than sequential. I suspect this may be an artefact of detecting resistance thresholds when the threshold is already reached for i2.

~ created pap1_figs_slimmed_50_unequal.Rmd
~ all changes marked with #4unequal

Seems that resistance is slower for mixtures in all replicates.
For the 500 runs at least.

I have been racking my brain to think if we can address this question using  data that have already been generated (and hence avoiding having to explain how we generated new data) . Rather than trying to explain it informally in an Email, then write it up, I went directly to step 2 and wrote the following as a draft paragraph for the ms…. Does it make sense???

“We can analyse a subset of the sensitivity analysis to investigate this specific question of whether single  use of a new insecticide is better than deploying it as a mixture with a pre-existing insecticide which already has some resistance present in the population.  We selected runs where the starting frequency of resistance to insecticide 2 was > 10 fold higher than frequency of resistance to insecticide 1. We then compared the time for insecticide 1 resistance to reach 50% when deployed on its own, compared to time for resistance to insecticide 1 to reach 50% when deployed as a mixture with the pre-existing insecticide. As in the full analysis, we compute the ratio mixture/single for each run and generate a classification tree to see what circumstances favour what policy and, in particular whether the ratio of resistance starting frequencies has a large impact”

I am assuming insecticide #1 is always deployed first in the sequential runs so the time to its resistance reaching 50% is already recorded. I am also assuming that in the mixture runs you have recorded  the time until resistance to each insecticide reaches 50% so you always have the data #1.

this is the important bit from Ian :
We selected runs where the starting frequency of resistance to insecticide 2 was > 10 fold higher than frequency of resistance to insecticide 1. We then compared the time for insecticide 1 resistance to reach 50% when deployed on its own, compared to time for resistance to insecticide 1 to reach 50% when deployed as a mixture with the pre-existing insecticide.

I think it would be easier to do this by new runs, which is close to what what I've done in pap1_figs_slimmed_50_unequal. (although I might want to change the freq of the 'old' insecticide from a constant 0.5 to a range e.g. 0.2 - 0.5, but then would).

Could I practically do what Ian has suggested ? 
How many runs are there where the start freq of i2 is > 10 * i1 ?
Where resist_start_1_div_2 < 0.1

see in paper1_results_figs.rmd

note my hi_div_lo was actually lo_div_hi, corrected in the figs_slimmed_50.rmd

resist_start_hi_div_lo <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])

@@

aha! had a thought. Can I highlight those runs in the killer plot where resist_start_1_div_2 < 0.1 ?

probably need to do it for the mix2 graph rather than mixadaptive.

1. tidy these notes

@@

### current state of play 23/3/16
1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
1. runModel2.r          : runs model for the passed scenarios
1. paper1_results_figs_slimmed50.Rmd : final figures for paper
